\tokenWord:\\a Xanalogical Transclusion and Micropayment System

### 2003-02-15 papers/hypertext2003/paper.tex
Changed title.
Spell-checked.

%
% Modification History
%
% 2003-February-6  Jason Rohrer
% Created.
%
% 2003-February-13 Jason Rohrer
% Finished initial draft.
% Spell-checked.
% Finished first round of proof reading.
%
% 2003-February-14 Jason Rohrer
% Made changes to match ACM SIG formatting guidelines.
%
% 2003-February-15 Jason Rohrer
% Added changes in response to Jim's comments.
%

@@

\newcommand{\tw}{token\_word}

@@

\title{\tw:  a Xanalogical Storage and Authoring System}

@@

\begin{abstract}
We describe \tw, a web-based xanalogical hypertext system.  
Our system supports almost every core feature associated with Xanadu, an architecture that has been actively discussed for nearly forty years.
Features include deep quotation, unbreakable references, two-way reference chasing, and frictionless micropayments that are backed by real-world currency.
To our knowledge, \tw \ is the first xanalogical system implementation to included all of these features.  Furthermore, \tw \  is the first xanalogical system of any kind to be accessible by a web browser with no additional software.
Previous attempts at full xanalogical implementations took many years and were ultimately never completed.
Given this historical backdrop, an additional contribution of this paper is in analyzing the development strategies that led to a relatively novel success.
A live \tw \  system is running at:

@@

\section{System Architecture}
We designed and built \tw, a full-featured xanalogical system, from scratch in only ten days, and our development team consisted of a single programmer.
To achieve this kind of rapid development, we relied heavily on modern technology and development techniques.
We do not view past Xanadu failures with disdain.
Those early attempts were simply ahead of their time:  the technology of the day was not ready for Xanadu.
The central technologies that made \tw \   development possible were the web, CGI, Perl, and PayPal \cite{paypal}.

Our system, though full-featured, was not designed to be ``the'' scalable xanalogical system that will carry us all into the future.
Instead, we focused on building a usable implementation of core xanalogical features for the purpose of experimentation.
Though scalability was not a design goal, \tw \  was built on top of technologies that have scaled well for other web-based projects \cite{Everything2, WikiWikiWeb, Wikipedia}.
Given these precedents, our architecture could potentially be scalled to support large document collections.
Section \ref{sec:Scalability} discusses these scalability issues in more detail.

\subsection{User Perspective}
In describing the \tw \  user experience, we are faced with the very issue that motivated \tw \  development in the first place:  descriptions in words and pictures do not give us a true sense of what it is like to \textit{use} a system.
Since this paper's focus is on a real-world implementation and not just an abstract description, we encourage readers to visit \texttt{http://hypertext.sf.net/token\_word} to try our system for themselves.

From the end-user's perspective, \tw \  has many sal\-ient features.
Users access \tw \  with a web browser, and they must create accounts and log in to use the system.  
The central elements of the system are documents, which can contain quotes of other documents.
Quote markers can be hidden or displayed, and quotes can be followed to see the context from which they were taken.
Each document also features a list of documents that quote it, so references can be followed in both directions.
The \tw \   document display interface can be seen in Figure \ref{fig:mainScreen}.

@@

Figures \ref{fig:extractQuote} through \ref{fig:docPreview} show a sequence of actions from a common usage scenareo:  extracting a quote, viewing the quote on the clipboard, creating a new document that uses the quote, and previewing the document.
After correcting the spelling mistake highlithed in Figure \ref{fig:docPreview}, the final step would be to submit the document.
Once submitted, the work becomes a full-fledged \tw \  document, making it readable and quotable by other users in the system.

Each user has a token balance, where one token is worth one character.
The current exchange rate is 1,000,000 tokens for \$1, and Paypal is integrated for token deposits and withdrawls.
When a reader views a document that contains as of yet unviewed text regions, tokens are transfered from the reader to the original authors of the regions.
Users only pay for each piece of text once---repeat viewings of the same text, in any context, are free.
\begin{figure}[t]
\centering
\epsfig{file=images/docCreate.eps, width=20pc}
\caption{The document creation interface. Quote 1 from the clipboard is inserted with the $<$q1$>$ tag.}
\label{fig:docCreate}
\end{figure}

\begin{figure}[t]
\centering
\epsfig{file=images/docPreview.eps, width=20pc}
\caption{The document preview interface.  Quotes are expanded, and possible spelling mistakes are highlighed.}
\label{fig:docPreview}
\end{figure}

Full text searches are supported across the entire document space, and search terms are highlighted when result documents are displayed.
Search results (and other lists in \tw) are sorted with most-quoted documents first, since frequently quoted documents are likely to contain the best information.
This ranking mechanism is similar to Google's PageRank algorithm \cite{Brin1998}.

\tw \  provides no explicit document linking mechanism.
However, since quote context can be obtained easily, brief quotes can be used to emulate links between documents.
For example, to link to a target document called ``frequently asked questions,'' the title words of the document can be quoted.
When quote markers are shown in the linking document, the words ``frequently asked questions'' will be marked, and clicking on the words will lead to the context of the quote (in other words, to the target document ``frequently asked questions'').
If the title words are not appropriate for use as a link anchor, another more relevant phrase from the body of the target document can be quoted to form a link.
For an example of quotes used as links, see Figure \ref{fig:quotesAsLinks}.

\begin{figure}[t]
\centering
\epsfig{file=images/quotesAsLinks.eps, width=20pc}
\caption{Brief portions of documents can be quoted to emulate links.  Here, the titles of two documents, ``frequently asked questions'' and ``\tw \  history,'' have been quoted.}
\label{fig:quotesAsLinks}
\end{figure}

@@

\subsubsection{Data Model}
Keeping with the Xanadu model, \tw \  separates documents from the text they contain.
There are two types of data objects in \tw:  documents and chunks.
A chunk contains a raw block of contiguous text, while a document is a list of references to text chunks.
Some of the referenced chunks in a document may have been created at the same time as the document (for new text in the document), while others may be part of quotes taken from other documents.

@@

\subsubsection{Development}
The entire \tw \   system is implemented as a single CGI script written in Perl, though the script is broken up into eight modules for organization purposes.
Perl's built-in support for regular expressions was useful for the rather complicated text processing done by \tw.
Many of these text processing operations can be expressed in succinct and elegant ways using Perl.
For example, highlighting possible misspellings in preview mode was a minor addition that required less than fifteen lines of code.

User interfaces can be layed out easily with HTML, and CGI processing is well-supported by existing Perl modules.
The CGI model requires a stateless action-response programming style, which makes it easy to focus on one feature at a time during development.
Perl-based CGI programming seems to be the modern method of choice for rapidly developing complex, multi-user applications.
Other systems that have used this development method successfully are discussed in Section \ref{sec:RelatedWork}.

Even though the full development cycle was short, we used an iterative process.
Features were added one at a time to \tw, and the system was maintained in a usable form at all times.
For example, we started with a system that managed text chunks, then added document management, then added quotes, and then added tokens transfers.
Final iterations added features such as spell checking and index-based searching.
From our understanding of previous xanalogical implementation efforts, this iterative process may have been key to our success.
The Xanadu model is complex and subtle, and it is easy to get bogged down by details without ever producing a working system.
Our system started out simple and grew in complexity, feature by feature.

@@

If our system did need to scale to support millions of users and billions of chunks and documents, several approaches would be possible.
First, we could use an existing database system as a backend to store chunks and documents.
This approach would deal with the inode limitations, and it might actually improve performance, since database queries would replace the directory lookups that are currently used to fetch chunks and documents.
A second approach would be to implement a simple application-level virtual filesystem for chunk and document storage.
With this approach, we could taylor the storage system to our specific needs, though its development would certainly be labor-intensive.

All filesystem operations in \tw \  are wrapped by a single module with only a handful of interface functions.
By modifying the implementation of these module functions, either of the scalable backends mentioned above could be added to \tw \  with no changes to the rest of the code.

@@

\section{Micropayment Issues}
With our micropyament system, we wanted to leverage PayPal's \cite{paypal} immense popularity and familiarity.
By handling real-world currency exclusively through PayPal, instead of through our own virtual money system, we can process payments in a secure and trustworthy fashion.

However, PayPal's currency model does not support the microdollar resolution that we need, so we cannot rely on this service alone for transfering tokens between users.
Even if PayPal did support high-resolution payments, it does not allow a third party to initiate transfers between account holders in a transparent and frictionless way (and for good reason, given the security issues).

To deal with these issues, we built an internal token balance system and used PayPal integration only for deposits and withdrawls of real-world currency.
With this configuration, a deposit is by no means frictionless (logging in to PayPal is required), but after the deposit has been made, token payments are completely transparent as the user accesses \tw \  documents.
Similarly, withdrawing tokens in exchange for real-world currency is not frictionless: 
PayPal does not support an automated payment mechanism, so withdrawl requests must be queued for manual processing by the \tw \  system administrator.
Whereas deposits take effect immediately (using PayPal's ``Instant Payment Notification'' service), withdrawn token refunds are delayed by the response of the human administrator.
From the administrator's point of view, the withdrawl queue can be used to increase security, since any suspicious activity can be investigated before real-world payments are made.

@@

\subsection{Balancing Risk}
Given that micropayments in \tw \  are mandatory and automatic, a major barrier to system adoption might exist:  trying \tw \  would require the deposit of real-world currency.
Even the main page displayed after login is a first-class \tw \  document, so a new user with no tokens would not even be able to read a description of \tw, let alone try any of its features.

One way of dealing with this adoption barrier would be to give each new user a small number of ``free'' tokens to start with (for example, at least enough tokens to read the main page).
This solution would allow new users to try the system with no monetary investment, but it carries an enormous risk for the \tw \  system administrators.
A small group of users could create many fake accounts and then use each account to read a large dummy document owned by one user.
This kind of scam could be used to consolodate the free starting tokens from many accounts into one account, and then a withdrawl could be made in exchange for a large amount of real-world currency.
Even if users are not trying to intentially scam the system, doling out free tokens poses considerable risk.
One user might write documents that are popular with new users (for example, a collection of how-to documents) and ammass a large number of free tokens over time---the withdrawl of these tokens could financially devastate the \tw \  administrators.

We have devised a risk-free mechanism that deals with the adoption barrier.  
Our approach uses two token types.
The first type of tokens, called \textit{trial} tokens, have no real-world value.
Each new account starts out with a small number of trial tokens (in our live system, each account starts with 50,000 trial tokens).
The second type of tokens, called \textit{real} tokens, can be exchanged for real-world currency.
The token transfer logic in \tw \  keeps track of token types:  trial tokens can only be transfered between users' trial balances, and real tokens can only be transfered between real balances.
Trial tokens are favored, and real tokens are only spent when a user has no trial tokens.

For example, suppose that the balances of three users are as shown in following table:
%suppose user $A$ has a trial:real token balance of 50:200, user $B$ has a balance of 0:500, and user $C$ has a balance of 10:0.
\begin{center}
\begin{tabular}{|l||r|r|}
\hline
User&Trial balance &Real balance\\
\hline
$A$&50 &200\\
\hline
$B$&0  &500\\
\hline
$C$&10 &0  \\
\hline
\end{tabular}
\end{center}
Suppose that user $A$ reads a new 100-chracter document that is owned by $B$, but that the document contains a 20-character quote by $C$.
User $A$ must pay 100 tokens to access this new content, with 80 going to $B$ and 20 going to $C$.
The transfer logic favors trial tokens, so $A$'s trial balance is exhausted first by transfering 50 tokens to $B$.
The remaining payment to $B$ is made with 30 real tokens, and 20 real tokens are also transfered to $C$.
The resulting balances are shown in the following table:
\begin{center}
\begin{tabular}{|l||r|r|}
\hline
User&Trial balance &Real balance\\
\hline
$A$&0 &150\\
\hline
$B$&50  &530\\
\hline
$C$&10 &20  \\
\hline
\end{tabular}
\end{center}

@@

This two-type scheme allows users to try \tw \  extensively without a monetary investment, and it also poses no financial risks to the administrators.
Note that though trial tokens have no real-world value, they still have value in \tw, since they can always be used to purchase new content.
Inside \tw, a user that amasses many trial tokens is just as wealthy as a user that has many real tokens.
A user that actively publishes popular content in \tw \  could effectively use the system forever without depositing any real-world money.

@@

In terms of implementation, \tw \  is closely related to a recent class of hypertext systems known as Wikis, which are often implemented as Perl-based CGI programs.
Wikis are reader-edited web sites, and web-based HTML forms are used as the interface for page authoring.
In many cases, Wikis are open for editing by the public at large.

@@

%\tw \  differs from a Wiki in its approach to hypertext, since a \tw \  quote can only be taken from an existing document.
%No mechanism is provided for refering to a document that does not yet exist.

@@

\subsection{Everything2}
Like \tw \  and Wikis, Everything2 \cite{Everything2} is a web-based hypertext system implemented using CGI and Perl.
Everything2 supports dangling links to pages that do not exist yet, but it handles these links differently than a Wiki does.
When a user follows a dangling link, a search is performed for the title words of the non-existant page.
Thus, dangling links are more fuctional in Everything2 from a reader's point of view.
Using this comparison point, we might conclude that Wikis favor authors, while Everything2 favors readers.
%As soon as a page is created that matches the title of a dangling link, that link points to the new page directly.

@@

\subsection{Commonalities and Differences}
All three systems discussed above share several features in common with \tw.
First, they are all web-based, collaborative information spaces.
Second, they all rely heavily on existing technology for their implementations.
Third, from what we can tell, they were developed relatively quickly.

@@

Of course, out of all of the collaborative information spaces discussed here, \tw \  is the only system that supports deep content reuse and frictionless micropayments.
Thus, these systems are closely related in terms of implementation, but their goals and end-user experiences are very different.

\section{Conclusion}
Thoughout this paper, we have discussed the strategies that made \tw \  development possible, including an iterative process and heavy use of existing technology.

@@

\tw \  is by no means an ideal xanalogical system, and many features are missing.
However, it is still the first system to implement the core features of the Xanadu architecture.
Though one would be hard-pressed to argue that the features included in \tw \  are not crucial, three important features are missing.
First, \tw \  does not explicitly support document versioning.
However, using the quotation primitive, a limited form of versioning can be emulated (a user can quote unchanged regions of the previous version when creating a new version).
Second, \tw \  does not support multi-server integration:  each installation of \tw \  operates in isolation, and quotes between documents at different sites are not possible.
We see multi-server integration as one of the weakpoints in the Xanadu architecture:  how can we ensure that references never break in a system comprised of multiple, separately owned servers?
Even if fault-tolerant, multi-server integration is possible, it is not essential for the end-user experience. 
Third, \tw \   does not support side-by-side document comparission with \textit{transpointing} lines, a classic Xanadu interface feature that Nelson's recent paper describes in detail \cite{Nelson1999}.
The transpointing interface would be difficult, if not impossible, to support in a web browser with no additional software.
However, a side-by-side document comparisson interface, using colored highlighting to indicate itentical regions, has been contemplated for \tw.
Implementing such an interface would be easy, though coming up with a highlighting scheme that works correctly in the general case (for example, without running out of readable colors) is quite difficult.

In recent personal conversations, Nelson has described a layered presentation model that could work in conjunction with a xanalogical storage system.
For example, a presentation layer could specify font sizes, links, or annotations, and multiple layers could be stacked when rendering a particular piece of content.
Given \tw's current data model and presentation logic, adding layers would not be difficult.
In fact, particular types of layers are already implemented:  highlights for misspellings, quote context, and hit search terms are applied to document text after it is assembled.
Highlights are achieved by inserting presentation tags (in this case, HTML font color tags) around particular character spans.
Many other presentation styles could be supported in this way by inserting other tag types.
Layers are an ideal feature to add in a future development iteration.

>>>>

%
% Modification History
%
% 2003-February-6  Jason Rohrer
% Created.
%
% 2003-February-13 Jason Rohrer
% Finished initial draft.
% Spell-checked.
% Finished first round of proof reading.
%
% 2003-February-14 Jason Rohrer
% Made changes to match ACM SIG formatting guidelines.
%
% 2003-February-15 Jason Rohrer
% Added changes in response to Jim's comments.
% Spell-checked.
%

@@

\newcommand{\tokenWord}{token\_word}

@@

\title{\tokenWord:\\a Xanalogical Transclusion and Micropayment System}

@@

\begin{abstract}
We describe \tokenWord, a web-based xanalogical hypertext system.  
Our system supports almost every core feature associated with Xanadu, an architecture that has been actively discussed for nearly forty years.
Features include deep quotation, unbreakable references, two-way reference chasing, and frictionless micropayments that are backed by real-world currency.
To our knowledge, \tokenWord \  is the first xanalogical system implementation to included all of these features.  Furthermore, \tokenWord \  is the first xanalogical system of any kind to be accessible by a web browser with no additional software.
Previous attempts at full xanalogical implementations took many years and were ultimately never completed.
Given this historical backdrop, an additional contribution of this paper is in analyzing the development strategies that led to a relatively novel success.
A live \tokenWord \  system is running at:

@@

\section{System Architecture}
We designed and built \tokenWord, a full-featured xanalogical system, from scratch in only ten days, and our development team consisted of a single programmer.
To achieve this kind of rapid development, we relied heavily on modern technology and development techniques.
We do not view past Xanadu failures with disdain.
Those early attempts were simply ahead of their time:  the technology of the day was not ready for Xanadu.
The central technologies that made \tokenWord \   development possible were the web, CGI, Perl, and PayPal \cite{paypal}.

Our system, though full-featured, was not designed to be ``the'' scalable xanalogical system that will carry us all into the future.
Instead, we focused on building a usable implementation of core xanalogical features for the purpose of experimentation.
Though scalability was not a design goal, \tokenWord \  was built on top of technologies that have scaled well for other web-based projects \cite{Everything2, WikiWikiWeb, Wikipedia}.
Given these precedents, our architecture could potentially be scaled to support large document collections.
Section \ref{sec:Scalability} discusses these scalability issues in more detail.

\subsection{User Perspective}
In describing the \tokenWord \  user experience, we are faced with the very issue that motivated \tokenWord \  development in the first place:  descriptions in words and pictures do not give us a true sense of what it is like to \textit{use} a system.
Since this paper's focus is on a real-world implementation and not just an abstract description, we encourage readers to visit \texttt{http://hypertext.sf.net/token\_word} to try our system for themselves.

From the end-user's perspective, \tokenWord \  has many sal\-ient features.
Users access \tokenWord \  with a web browser, and they must create accounts and log in to use the system.  
The central elements of the system are documents, which can contain quotes of other documents.
Quote markers can be hidden or displayed, and quotes can be followed to see the context from which they were taken.
Each document also features a list of documents that quote it, so references can be followed in both directions.
The \tokenWord \  document display interface can be seen in Figure \ref{fig:mainScreen}.

@@

Figures \ref{fig:extractQuote} through \ref{fig:docPreview} show a sequence of actions from a common usage scenario:  extracting a quote, viewing the quote on the clipboard, creating a new document that uses the quote, and previewing the document.
After correcting the spelling mistake highlighted in Figure \ref{fig:docPreview}, the final step would be to submit the document.
Once submitted, the work becomes a full-fledged \tokenWord \  document, making it readable and quotable by other users in the system.

Each user has a token balance, where one token is worth one character.
The current exchange rate is 1,000,000 tokens for \$1, and PayPal is integrated for token deposits and withdrawls.
When a reader views a document that contains as of yet unviewed text regions, tokens are transfered from the reader to the original authors of the regions.
Users only pay for each piece of text once---repeat viewings of the same text, in any context, are free.
\begin{figure}[t]
\centering
\epsfig{file=images/docCreate.eps, width=20pc}
\caption{The document creation interface. Quote 1 from the clipboard is inserted with the $<$q1$>$ tag.}
\label{fig:docCreate}
\end{figure}

\begin{figure}[t]
\centering
\epsfig{file=images/docPreview.eps, width=20pc}
\caption{The document preview interface.  Quotes are expanded, and possible spelling mistakes are highlighted.}
\label{fig:docPreview}
\end{figure}

Full text searches are supported across the entire document space, and search terms are highlighted when result documents are displayed.
Search results (and other lists in \tokenWord) are sorted with most-quoted documents first, since frequently quoted documents are likely to contain the best information.
This ranking mechanism is similar to Google's PageRank algorithm \cite{Brin1998}.

\tokenWord \  provides no explicit document linking mechanism.
However, since quote context can be obtained easily, brief quotes can be used to emulate links between documents.
For example, to link to a target document called ``frequently asked questions,'' the title words of the document can be quoted.
When quote markers are shown in the linking document, the words ``frequently asked questions'' will be marked, and clicking on the words will lead to the context of the quote (in other words, to the target document ``frequently asked questions'').
If the title words are not appropriate for use as a link anchor, another more relevant phrase from the body of the target document can be quoted to form a link.
For an example of quotes used as links, see Figure \ref{fig:quotesAsLinks}.

\begin{figure}[t]
\centering
\epsfig{file=images/quotesAsLinks.eps, width=20pc}
\caption{Brief portions of documents can be quoted to emulate links.  Here, the titles of two documents, ``frequently asked questions'' and ``\tokenWord \  history,'' have been quoted.}
\label{fig:quotesAsLinks}
\end{figure}

@@

\subsubsection{Data Model}
Keeping with the Xanadu model, \tokenWord \  separates documents from the text they contain.
There are two types of data objects in \tokenWord:  documents and chunks.
A chunk contains a raw block of contiguous text, while a document is a list of references to text chunks.
Some of the referenced chunks in a document may have been created at the same time as the document (for new text in the document), while others may be part of quotes taken from other documents.

@@

\subsubsection{Development}
The entire \tokenWord \   system is implemented as a single CGI script written in Perl, though the script is broken up into eight modules for organization purposes.
Perl's built-in support for regular expressions was useful for the rather complicated text processing done by \tokenWord.
Many of these text processing operations can be expressed in succinct and elegant ways using Perl.
For example, highlighting possible misspellings in preview mode was a minor addition that required less than fifteen lines of code.

User interfaces can be laid out easily with HTML, and CGI processing is well-supported by existing Perl modules.
The CGI model requires a stateless action-response programming style, which makes it easy to focus on one feature at a time during development.
Perl-based CGI programming seems to be the modern method of choice for rapidly developing complex, multi-user applications.
Other systems that have used this development method successfully are discussed in Section \ref{sec:RelatedWork}.

Even though the full development cycle was short, we used an iterative process.
Features were added one at a time to \tokenWord, and the system was maintained in a usable form at all times.
For example, we started with a system that managed text chunks, then added document management, then added quotes, and then added tokens transfers.
Final iterations added features such as spell checking and index-based searching.
From our understanding of previous xanalogical implementation efforts, this iterative process may have been key to our success.
The Xanadu model is complex and subtle, and it is easy to get bogged down by details without ever producing a working system.
Our system started out simple and grew in complexity, feature by feature.

@@

If our system did need to scale to support millions of users and billions of chunks and documents, several approaches would be possible.
First, we could use an existing database system as a backend to store chunks and documents.
This approach would deal with the inode limitations, and it might actually improve performance, since database queries would replace the directory lookups that are currently used to fetch chunks and documents.
A second approach would be to implement a simple application-level virtual filesystem for chunk and document storage.
With this approach, we could tailor the storage system to our specific needs, though its development would certainly be labor-intensive.

All filesystem operations in \tokenWord \  are wrapped by a single module with only a handful of interface functions.
By modifying the implementation of these module functions, either of the scalable backends mentioned above could be added to \tokenWord \  with no changes to the rest of the code.

@@

\section{Micropayment Issues}
With our micropayment system, we wanted to leverage PayPal's \cite{paypal} immense popularity and familiarity.
By handling real-world currency exclusively through PayPal, instead of through our own virtual money system, we can process payments in a secure and trustworthy fashion.

However, PayPal's currency model does not support the microdollar resolution that we need, so we cannot rely on this service alone for transferring tokens between users.
Even if PayPal did support high-resolution payments, it does not allow a third party to initiate transfers between account holders in a transparent and frictionless way (and for good reason, given the security issues).

To deal with these issues, we built an internal token balance system and used PayPal integration only for deposits and withdrawls of real-world currency.
With this configuration, a deposit is by no means frictionless (logging in to PayPal is required), but after the deposit has been made, token payments are completely transparent as the user accesses \tokenWord \  documents.
Similarly, withdrawing tokens in exchange for real-world currency is not frictionless: 
PayPal does not support an automated payment mechanism, so withdrawl requests must be queued for manual processing by the \tokenWord \  system administrator.
Whereas deposits take effect immediately (using PayPal's ``Instant Payment Notification'' service), withdrawn token refunds are delayed by the response of the human administrator.
From the administrator's point of view, the withdrawl queue can be used to increase security, since any suspicious activity can be investigated before real-world payments are made.

@@

\subsection{Balancing Risk}
Given that micropayments in \tokenWord \  are mandatory and automatic, a major barrier to system adoption might exist:  trying \tokenWord \  would require the deposit of real-world currency.
Even the main page displayed after login is a first-class \tokenWord \  document, so a new user with no tokens would not even be able to read a description of \tokenWord, let alone try any of its features.

One way of dealing with this adoption barrier would be to give each new user a small number of ``free'' tokens to start with (for example, at least enough tokens to read the main page).
This solution would allow new users to try the system with no monetary investment, but it carries an enormous risk for the \tokenWord \  system administrators.
A small group of users could create many fake accounts and then use each account to read a large dummy document owned by one user.
This kind of scam could be used to consolidate the free starting tokens from many accounts into one account, and then a withdrawl could be made in exchange for a large amount of real-world currency.
Even if users are not trying to intentionally scam the system, doling out free tokens poses considerable risk.
One user might write documents that are popular with new users (for example, a collection of how-to documents) and amass a large number of free tokens over time---the withdrawl of these tokens could financially devastate the \tokenWord \  administrators.

We have devised a risk-free mechanism that deals with the adoption barrier.  
Our approach uses two token types.
The first type of tokens, called \textit{trial} tokens, have no real-world value.
Each new account starts out with a small number of trial tokens (in our live system, each account starts with 50,000 trial tokens).
The second type of tokens, called \textit{real} tokens, can be exchanged for real-world currency.
The token transfer logic in \tokenWord \  keeps track of token types:  trial tokens can only be transfered between users' trial balances, and real tokens can only be transfered between real balances.
Trial tokens are favored, and real tokens are only spent when a user has no trial tokens.

For example, suppose that the balances of three users are as shown in following table:
%suppose user $A$ has a trial:real token balance of 50:200, user $B$ has a balance of 0:500, and user $C$ has a balance of 10:0.
\begin{center}
\begin{tabular}{|l||r|r|}
\hline
User&Trial balance &Real balance\\
\hline
$A$&50 &200\\
\hline
$B$&0  &500\\
\hline
$C$&10 &0  \\
\hline
\end{tabular}
\end{center}
Suppose that user $A$ reads a new 100-character document that is owned by $B$, but that the document contains a 20-character quote by $C$.
User $A$ must pay 100 tokens to access this new content, with 80 going to $B$ and 20 going to $C$.
The transfer logic favors trial tokens, so $A$'s trial balance is exhausted first by transferring 50 tokens to $B$.
The remaining payment to $B$ is made with 30 real tokens, and 20 real tokens are also transfered to $C$.
The resulting balances are shown in the following table:
\begin{center}
\begin{tabular}{|l||r|r|}
\hline
User&Trial balance &Real balance\\
\hline
$A$&0 &150\\
\hline
$B$&50  &530\\
\hline
$C$&10 &20  \\
\hline
\end{tabular}
\end{center}

@@

This two-type scheme allows users to try \tokenWord \  extensively without a monetary investment, and it also poses no financial risks to the administrators.
Note that though trial tokens have no real-world value, they still have value in \tokenWord, since they can always be used to purchase new content.
Inside \tokenWord, a user that amasses many trial tokens is just as wealthy as a user that has many real tokens.
A user that actively publishes popular content in \tokenWord \  could effectively use the system forever without depositing any real-world money.

@@

In terms of implementation, \tokenWord \  is closely related to a recent class of hypertext systems known as Wikis, which are often implemented as Perl-based CGI programs.
Wikis are reader-edited web sites, and web-based HTML forms are used as the interface for page authoring.
In many cases, Wikis are open for editing by the public at large.

@@

%\tokenWord \  differs from a Wiki in its approach to hypertext, since a \tokenWord \  quote can only be taken from an existing document.
%No mechanism is provided for referring to a document that does not yet exist.

@@

\subsection{Everything2}
Like \tokenWord \  and Wikis, Everything2 \cite{Everything2} is a web-based hypertext system implemented using CGI and Perl.
Everything2 supports dangling links to pages that do not exist yet, but it handles these links differently than a Wiki does.
When a user follows a dangling link, a search is performed for the title words of the non-existent page.
Thus, dangling links are more functional in Everything2 from a reader's point of view.
Using this comparison point, we might conclude that Wikis favor authors, while Everything2 favors readers.
%As soon as a page is created that matches the title of a dangling link, that link points to the new page directly.

@@

\subsection{Commonalities and Differences}
All three systems discussed above share several features in common with \tokenWord.
First, they are all web-based, collaborative information spaces.
Second, they all rely heavily on existing technology for their implementations.
Third, from what we can tell, they were developed relatively quickly.

@@

Of course, out of all of the collaborative information spaces discussed here, \tokenWord \  is the only system that supports deep content reuse and frictionless micropayments.
Thus, these systems are closely related in terms of implementation, but their goals and end-user experiences are very different.

\section{Conclusion}
Throughout this paper, we have discussed the strategies that made \tokenWord \  development possible, including an iterative process and heavy use of existing technology.

@@

\tokenWord \  is by no means an ideal xanalogical system, and many features are missing.
However, it is still the first system to implement the core features of the Xanadu architecture.
Though one would be hard-pressed to argue that the features included in \tokenWord \  are not crucial, three important features are missing.
First, \tokenWord \  does not explicitly support document versioning.
However, using the quotation primitive, a limited form of versioning can be emulated (a user can quote unchanged regions of the previous version when creating a new version).
Second, \tokenWord \  does not support multi-server integration:  each installation of \tokenWord \  operates in isolation, and quotes between documents at different sites are not possible.
We see multi-server integration as one of the weakpoints in the Xanadu architecture:  how can we ensure that references never break in a system comprised of multiple, separately owned servers?
Even if fault-tolerant, multi-server integration is possible, it is not essential for the end-user experience. 
Third, \tokenWord \   does not support side-by-side document comparison with \textit{transpointing} lines, a classic Xanadu interface feature that Nelson's recent paper describes in detail \cite{Nelson1999}.
The transpointing interface would be difficult, if not impossible, to support in a web browser with no additional software.
However, a side-by-side document comparison interface, using colored highlighting to indicate identical regions, has been contemplated for \tokenWord.
Implementing such an interface would be easy, though coming up with a highlighting scheme that works correctly in the general case (for example, without running out of readable colors) is quite difficult.

In recent personal conversations, Nelson has described a layered presentation model that could work in conjunction with a xanalogical storage system.
For example, a presentation layer could specify font sizes, links, or annotations, and multiple layers could be stacked when rendering a particular piece of content.
Given \tokenWord's current data model and presentation logic, adding layers would not be difficult.
In fact, particular types of layers are already implemented:  highlights for misspellings, quote context, and hit search terms are applied to document text after it is assembled.
Highlights are achieved by inserting presentation tags (in this case, HTML font color tags) around particular character spans.
Many other presentation styles could be supported in this way by inserting other tag types.
Layers are an ideal feature to add in a future development iteration.
